\documentclass[lettersize,journal]{IEEEtran}

% ── Packages ──────────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage[table]{xcolor}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
% ── Placeholder colour ────────────────────────────────────────────────────
\newcommand{\placeholder}[1]{\textcolor{gray}{\textit{[#1]}}}

% ── Math shorthands ───────────────────────────────────────────────────────
% Preference vector and transformation matrix notation
\newcommand{\xvec}{\boldsymbol{x}}
\newcommand{\yvec}{\boldsymbol{y}}
\newcommand{\Ku}{K_{u}}
\newcommand{\Kuinv}{K_{u}^{-1}}

% ── Document ──────────────────────────────────────────────────────────────
\begin{document}

\title{Privacy-Preserving Intelligent Mobility for Individuals with
       Autism Spectrum Disorder}

\author{Ahmed Aouab Admou, Yahya Misbahi, Titouan Mathieu,
        Nawfal El-Hammadi, and Toumany Doumbouya%
\thanks{The authors are with the National Institute of Applied Sciences,
Toulouse, France.}
\thanks{Manuscript prepared for the Privacy and Intelligent Robotics
project, 2025--2026.}}

\markboth{Privacy-Preserving Intelligent Mobility for ASD}%
{Admou \MakeLowercase{\textit{et al.}}: Privacy-Preserving Intelligent Mobility for ASD}

\maketitle

% ─────────────────────────────────────────────────────────────────────────
\begin{abstract}
Smart mobility systems increasingly use artificial intelligence and
edge sensing to optimize urban transport, but they rarely account for
the sensory and behavioral needs of people with Autism Spectrum
Disorder (ASD). Existing sensory-aware navigation tools provide useful
personalization, yet they often rely on centralized storage or limited
privacy mechanisms for highly sensitive mobility and health-related
data. This study aimed to design and implement a privacy-preserving
intelligent mobility prototype for ASD users based on a three-tier
Cloud--MEC--Device architecture. The proposed system combines an
encrypted and matrix-obfuscated ASD profile vector, on-device
personalized inference, optional smartwatch-derived physiological
signals such as BPM, regional MEC context enrichment, and personalized
hierarchical federated learning. In the implemented
prototype, the phone-side model is decomposed as
$\mathbf{w}_i=\mathbf{w}_s+\mathbf{v}_i$, where only the shared
component $\mathbf{w}_s$ is federated and the personal component
$\mathbf{v}_i$ remains local. MEC nodes aggregate regional shared
updates and provide situational inputs such as crowding, delay, and
noise, while the cloud coordinates global shared model initialization.
Prototype validation using synthetic and manual demonstration data
shows that the web app performs local AI inference, shared updates are
federated through MEC nodes, personal components remain local, and the
cloud receives only regional shared model metadata. These results
demonstrate the technical feasibility of combining ASD-aware mobility
support with privacy-by-design learning.
\end{abstract}

\begin{IEEEkeywords}
Autism Spectrum Disorder, smart mobility, personalized federated
learning, mobile edge computing, privacy-by-design, sensory-aware
navigation.
\end{IEEEkeywords}

% ─────────────────────────────────────────────────────────────────────────
\section{Introduction}

\IEEEPARstart{S}{mart} city mobility systems increasingly rely on pervasive sensing,
context-aware services, and artificial intelligence to optimize urban
transport. These systems usually prioritize travel time, congestion, or
network efficiency, but they rarely model the sensory accessibility of
routes. For individuals with Autism Spectrum Disorder (ASD), mobility
is shaped not only by distance but also by noise, light, crowd density,
unexpected delays, and the availability of calm recovery spaces.
Exposure to these stimuli can produce anxiety, sensory overload, or
route avoidance, making conventional navigation insufficient for safe
and autonomous travel~\cite{maclennan2023}.

Several researchers have investigated sensory-aware and personalized
urban navigation. Cena et al.~\cite{cena2023} evaluated a map-based
application that indexes places according to sensory characteristics and
enables route filtering from an individual user profile. Other
assistive mapping approaches have explored crowdmapping and inclusive
accessibility information. In parallel, privacy-preserving learning has
been studied for sensitive medical and mobility contexts. Federated
learning has been proposed to avoid centralized training data
collection, while differential privacy and edge computing have been
used to reduce the exposure of personal records~\cite{adnan2022,
alwarafy2020}. Recent reviews of Mobility-as-a-Service platforms have
also highlighted centralized cloud storage as a major privacy weakness
for mobility data~\cite{garroussi2025}.

However, previous work has not fully addressed the combination of
ASD-specific personalization, real-time regional mobility context, and
end-to-end privacy preservation. Existing systems often personalize
navigation by storing sensitive user profiles or by relying on
centralized service providers. They also rarely distinguish between
shared population-level learning and strictly personal adaptation. As a
result, it remains unclear how an ASD-aware mobility system can provide
personalized discomfort prediction while preventing raw sensory
profiles, physiological data, and personal model components from
leaving the user's device.

This study aimed to design and implement a privacy-preserving
intelligent mobility prototype for ASD users. The system represents
sensory sensitivities as a normalized profile vector, protects this
vector through user-specific matrix obfuscation and encryption, and
combines it locally with contextual and physiological features for
discomfort prediction. These physiological features may be entered
manually or obtained from a connected smartwatch, for example through
BPM and other wearable health signals. The learning architecture follows a
Cloud--MEC--Device hierarchy: the phone performs local inference and
personalized training, MEC nodes aggregate regional shared model
updates and provide situational context, and the cloud coordinates
global shared model initialization.

The main contribution is a fully functional prototype that integrates
personalized federated learning into this architecture. The phone model
is decomposed into a shared component and a private personal component,
so that $\mathbf{w}_i=\mathbf{w}_s+\mathbf{v}_i$. Only
$\mathbf{w}_s$ is sent to the MEC; $\mathbf{v}_i$ remains local. The
prototype also demonstrates MEC-based context enrichment, encrypted
profile synchronization, regional and global shared model endpoints,
and an operational web interface that runs local AI inference. The rest
of the paper reviews related work, presents the mathematical and
privacy framework, describes the implementation stack, and reports
prototype validation results obtained from synthetic and manual test
data.

% ─────────────────────────────────────────────────────────────────────────
% ─────────────────────────────────────────────────────────────────────────
\section{Related Work}
\label{sec:background}

\subsection{Inclusive Mobility and the Privacy Paradox}
Smart city mobility research increasingly uses sensing, artificial
intelligence, and real-time transport data to optimize urban flow.
However, inclusive mobility systems often still assume neurotypical
sensory processing. For many people with Autism Spectrum Disorder
(ASD), public space is experienced as a dense sensory environment in
which noise, light, crowding, smell, heat, route disruption, and lack of
predictable recovery spaces can directly affect autonomy and safety
~\cite{maclennan2023}. Personalized assistance therefore requires
fine-grained data such as sensory preferences, physiological stress
signals, behavioral indicators, and location context. This creates a
privacy paradox: the same information needed to adapt mobility support
can reveal health status, daily routines, and vulnerable places if it is
centralized or shared with third parties.
The relevant sensory barriers are multidimensional: auditory overload
from sirens or announcements, visual discomfort from flickering or
reflective light, social stress from dense crowds, olfactory triggers,
spatial disorientation in complex stations, and thermal discomfort in
poorly ventilated transit environments. Prior work also emphasizes the
importance of predictable routes and recovery spaces, since unexpected
events such as a cancelled bus or a closed passage can be more
disruptive than a longer travel time~\cite{maclennan2023,beck2025}.

\subsection{Sensory-Aware Navigation Systems}
Existing sensory-aware systems show that neurodivergent mobility cannot
be reduced to shortest-path routing. PIUMA indexes places according to
sensory features and lets users filter routes using an individual
profile~\cite{cena2023}. Related crowdmapping approaches such as
mPASS, Maps4all, and CARES extend accessibility information by
collecting user reports about architectural or sensory barriers.
Together, these systems demonstrate the value of explicit sensory
information, visual route interfaces, and calm-space identification.
Their limitations are also consistent across the literature: coverage
depends heavily on community contributions, real-time events such as
delays or crowding are difficult to capture, and municipal accessibility
frameworks still prioritize physical disability over cognitive and
sensory accessibility~\cite{garroussi2025}.
PIUMA is particularly relevant because it combines three ideas that also
shape our prototype: a user-provided sensory profile, crowdmapped
ratings of sensory points of interest, and assistive routing based on
the user's current internal state. Earlier systems such as mPASS focus
more strongly on physical barriers, while Maps4all and CARES show that
community reporting and social participation are important but remain
insufficient when real-time regional context is missing.

\subsection{Privacy-Preserving AI for Mobility}
Most Mobility-as-a-Service platforms rely on centralized cloud
architectures, which expose sensitive records during transmission,
storage, analytics, or partner sharing~\cite{garroussi2025}. Standard
consent mechanisms may also be inaccessible to neurodivergent users
when they rely on legal language, long forms, or interface patterns that
reduce genuine understanding~\cite{beck2025}. Privacy-by-design
approaches therefore move computation closer to the user and minimize
what leaves the device. Multi-access Edge Computing (MEC) reduces
latency and limits raw-data exposure by processing regional context
near users~\cite{alwarafy2020}. Federated Learning (FL) trains models
locally and exchanges model updates rather than raw records. Differential
privacy (DP), usually implemented through clipping and calibrated noise,
can additionally harden data against inference attacks.

\begin{table}[!t]
\caption{Related Work Synthesis and Remaining Gap\label{tab:related_work_synthesis}}
\centering
\scriptsize
\renewcommand{\arraystretch}{1.2}
\begin{tabularx}{\columnwidth}{
>{\raggedright\bfseries}p{1.7cm}
>{\raggedright}X
>{\raggedright\arraybackslash}X}
\toprule
\rowcolor{gray!15}
\textbf{Area} & \textbf{Main Contribution} & \textbf{Remaining Gap} \\
\midrule
Sensory maps &
Expose noise, crowding, and other sensory attributes &
Often rely on static or crowdmapped data with limited real-time context \\
\midrule
MaaS privacy &
Identifies centralized storage and sharing risks &
Rarely addresses ASD-specific profile sensitivity and consent needs \\
\midrule
MEC and FL &
Moves computation closer to users and avoids raw-data upload &
Usually aggregates a single model without separating personal adaptation \\
\midrule
Accessible consent &
Improves readability and user autonomy &
Must be integrated into the technical privacy architecture \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Positioning of This Work}
The reviewed literature motivates a system that combines sensory-aware
mobility support, accessible privacy practices, and decentralized AI.
The gap addressed in this work is the lack of an integrated prototype
that separates regional knowledge from private personal adaptation. Our
architecture combines encryption, obfuscation, MEC-based situational
context, and personalized hierarchical Federated Learning.
% ---------------------------------------------------------------
\section{System Architecture and Mathematical Framework}
% ---------------------------------------------------------------
\label{sec:architecture}

\subsection{User Application and Local Data Flow}

The entry point of the system is a mobile application designed to help
an ASD user anticipate and reduce mobility-related discomfort. At first
launch, the user completes a structured sensory and behavioral survey.
The questionnaire asks about sensitivities such as noise, light, crowd
density, route disruption, transition difficulty, need for isolation,
and preferred recovery conditions. The answers are normalized locally
into an ASD profile vector $\mathbf{x}$, which becomes the stable
personal baseline used by the local model. This vector is not sent in
plaintext to the MEC or the cloud.

During daily use, the application performs check-ins. A check-in
captures the user's current state, for example fatigue, stress,
breathing difficulty, communication load, and perceived environmental
pressure. The app can also be connected to a smartwatch or wearable
sensor, with user permission, to read BPM, heart-rate variability,
activity level, sleep-related indicators, or other physiological
signals. These data are treated as local phone-side inputs. If no
wearable is connected, the app continues to work with manual check-in
answers, route, sound and light context, and MEC-provided situational information.

The purpose of the application is therefore twofold. First, it provides
real-time local inference: the phone combines the ASD profile, current
check-in, optional wearable data, and regional context to estimate a
discomfort score before or during a mobility episode. Second, it
supports privacy-preserving personalization: user feedback and local
observations update the on-device model, but raw survey answers,
physiological streams, check-in details, and the private personalization
component remain on the device. Only protected shared model information
is later sent toward the MEC layer for regional learning.

\subsection{Overview of the Three-Tier Hierarchical Architecture}

The proposed system is organized into three computational tiers that
interact bidirectionally: the user device (smartphone), Multi-access
Edge Computing (MEC) nodes deployed at transport infrastructure
locations, and a central cloud server. This hierarchy is denoted
Cloud~$\leftrightarrow$~MEC~$\leftrightarrow$~Device. Each tier includes
a machine learning model whose architecture is shared between the three
tiers, but each tier is responsible for a distinct scope of learning and
inference.
The device tier performs real-time, personalized discomfort prediction and local
model fine-tuning using data that never leaves the hardware. The MEC
tier aggregates model weights from multiple devices within a geographic
region and specializes the shared model to local environmental patterns
such as infrastructure layout, typical crowd dynamics, and regional
climate. The cloud tier aggregates updates from all MEC nodes to
maintain a global model capturing broad ASD mobility behaviors across
the entire user population, and distributes generalized model
initializations back down to MEC nodes and devices. No raw user data
propagates beyond the device at any point in this pipeline.

The hierarchy also separates time scales. The phone reacts at the
shortest time scale, because discomfort prediction must occur during a
walk, transit change, or station arrival. MEC nodes operate at an
intermediate time scale, updating regional shared models and situational
context as local transport conditions evolve. The cloud operates at the
longest time scale, learning general initialization weights across
regions and users. This separation is important for ASD mobility because
the system must combine stable individual preferences with fast changes
in the environment, such as a delayed metro line, a crowded platform, or
unexpected construction noise.

Operationally, the device owns the complete personal context. It
combines the recovered ASD survey profile, current check-in answers,
optional wearable signals, and MEC-provided context into the local
feature vector. In a full mobile deployment, the phone can also connect
to a smartwatch or wearable sensor to read BPM, heart-rate variability,
sleep quality, activity level, or similar physiological indicators. The
MEC owns regional context and regional aggregation only; it does not
need to know which sensory dimensions are most sensitive for a specific
user or receive smartwatch raw streams. The cloud owns global
coordination and encrypted synchronization metadata. This division makes
the privacy boundary explicit and avoids treating the cloud as a trusted
holder of raw records.

\subsection{Threat Model and Privacy Boundary}

The system assumes an honest-but-curious cloud backend and honest-but-
curious MEC operators: these entities execute the protocol but may
attempt to infer sensitive information from the data they receive. The
network is assumed to be observable by an external adversary, so
transport encryption is required for deployment. The database may also
be compromised, motivating encryption of stored profile material and
obfuscation of profile vectors prior to persistence. Malicious clients
that deliberately submit poisoned model updates are considered a future
robustness problem and are outside the current prototype scope.

\begin{table}[!t]
    \centering
    \caption{Privacy Boundary of the Cloud--MEC--Device Architecture\label{tab:privacy_boundary}}
    \scriptsize
    \renewcommand{\arraystretch}{1.25}
    \begin{tabularx}{\columnwidth}{
        >{\raggedright\bfseries}p{1.5cm}
        >{\raggedright}X
        >{\raggedright\arraybackslash}X}
        \toprule
        \rowcolor{gray!15}
        \textbf{Layer} & \textbf{Visible Data} & \textbf{Hidden Data} \\
        \midrule
        Phone &
        ASD profile, raw check-in context, physiological signals,
        $\mathbf{v}_i$, and effective model $\mathbf{w}_i$ &
        Other users' profiles and model states \\
        \midrule
        MEC &
        Situational context, pseudonymous risk scores, shared
        updates $\mathbf{w}_s$, sample counts, regional aggregate model &
        ASD profile vector, physiological raw data, local dataset,
        personal component $\mathbf{v}_i$ \\
        \midrule
        Cloud &
        Encrypted profile records, derived check-in metrics, regional
        shared models, global shared model metadata &
        Raw profile vector, raw check-in payload, raw physiology,
        personal component $\mathbf{v}_i$ \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{ASD Profile Representation and Input Feature Engineering}

Each user's sensory profile is represented as a normalized vector
\begin{equation}
    \mathbf{x} = (x_1, x_2, \dots, x_n)^\top \in [0,1]^n
\end{equation}
where each component $x_i$ encodes a scalar sensitivity score for a
distinct sensory or behavioral dimension, including noise tolerance,
light sensitivity, crowd density tolerance, susceptibility to routine
disruption, and transition difficulty. This vector is derived from a
structured onboarding questionnaire administered at first launch. The
vector $\mathbf{x}$ is computed locally and persists only on the device
in encrypted form, as described in Section~\ref{sec:crypto}.

At inference time, the profile vector is combined with a real-time
contextual feature vector
\begin{equation}
    \mathbf{c} = (c_1, c_2, \dots, c_m)^\top \in \mathbb{R}^m
\end{equation}
where contextual features include ambient noise level, estimated crowd
density at the current location, weather conditions, heart rate acquired
from a paired smartwatch or wearable device, time of day, and current
transit mode.
The composite input to the on-device model is the concatenated feature
vector
\begin{equation}
    \mathbf{z} = [\mathbf{x}^\top \mid \mathbf{c}^\top]^\top
    \in \mathbb{R}^{n+m}
\end{equation}
The on-device model $f_{\boldsymbol{\theta}}$ maps $\mathbf{z}$ to a
predicted discomfort score
\begin{equation}
    d = f_{\boldsymbol{\theta}}(\mathbf{z}) \in [0, 1]
\end{equation}
where $d$ represents the estimated probability of an imminent sensory
overload or stress episode. The model architecture is a compact
feedforward neural network composed of a fully connected layer, a ReLU
activation, and a final classification layer, selected to remain
suitable for local simulation and on-device execution.

In the implemented prototype, the feature groups are explicitly
separated by origin. User-derived features include the ASD sensory
profile, current energy, breathing state, communication difficulty, and
need for isolation. Physiological features include BPM, heart rate,
heart-rate variability when available, sleep quality, activity level,
and wearable availability. These values can come from manual check-in
fields or from a connected smartwatch, depending on the device
permissions granted by the user. Local environmental features include perceived noise,
light, crowd density, routine disruption, transition difficulty, social
load, and calm-space availability. MEC situational features include
regional signals such as metro delay, transport incident state,
platform crowding, construction noise, and quiet-area availability.
Observable support signals include pacing, increased stimming, and
shutdown indicators. All these features are fused locally on the phone;
the MEC supplies only regional context and never receives the complete
personal feature vector.

The feature design is intentionally modular. Survey-derived features
change slowly and describe stable sensitivities; physiological and
check-in features change at session time; MEC situational features
change with the transport environment. Keeping these groups explicit
allows the prototype to support partial sensing. For instance, if a
smartwatch BPM or heart-rate signal is unavailable, the local model can
still combine the ASD profile, self-reported state, and MEC context.
Conversely, when a watch is paired, the physiological stream can enrich
the prediction without leaving the device. This design also supports
future ablation studies, because each feature group can be removed or
perturbed to measure its contribution to discomfort prediction.

\subsection{Personalized On-Device Federated Learning}

To avoid forcing every user in a region to share a single identical
model, the phone model is decomposed into a shared component and a
personal adaptation component. In this notation, $\mathbf{w}$ denotes a
vector of model weights, or trainable parameters, used by the local
discomfort prediction model. For user $i$, the effective inference model
weights are
\begin{equation}
    \mathbf{w}_i = \mathbf{w}_s + \mathbf{v}_i
\end{equation}
where $\mathbf{w}_s$ is the shared regional weight vector received from
the nearest MEC node and $\mathbf{v}_i$ is the private personal
adaptation weight vector stored only on the user's phone. For a new
user, $\mathbf{v}_i$ is initialized to the zero vector, so the first
effective model is simply the MEC initialization.

During local training, the loss is computed on the effective model
$\mathbf{w}_s+\mathbf{v}_i$, but the two components are updated with
separate learning rates:
\begin{equation}
    \mathbf{w}_s \leftarrow \mathbf{w}_s
    - \eta_s \nabla \mathcal{L}_i(\mathbf{w}_s + \mathbf{v}_i)
\end{equation}
\begin{equation}
    \mathbf{v}_i \leftarrow \mathbf{v}_i
    - \eta_p \nabla \mathcal{L}_i(\mathbf{w}_s + \mathbf{v}_i)
\end{equation}
where $\eta_s$ controls how quickly the shared component adapts and
$\eta_p$ controls user-specific personalization. This design lets the
phone benefit from regional learning while preserving a private
adaptation vector that captures idiosyncratic sensory reactions,
feedback history, and behavioral patterns.

Only the updated shared component $\mathbf{w}_s$ is transmitted to the
MEC node. The personal component $\mathbf{v}_i$ is never serialized,
published, or stored outside the phone. The main privacy boundary
therefore comes from local data processing, non-transmission of raw
records, encrypted profile storage, and the strict separation of
$\mathbf{v}_i$. As an optional hardening layer, the shared update can be
clipped and noised before transmission:
\begin{equation}
    \tilde{\mathbf{w}}_{s,i} =
    \mathrm{clip}\!\left(\mathbf{w}_{s,i},\,C\right)
    + \mathcal{N}\!\left(\mathbf{0},\,\sigma^2\mathbf{I}\right)
\end{equation}
The transmitted payload therefore contains only
$\tilde{\mathbf{w}}_{s,i}$, the local sample count $n_i$, and coarse
training metadata. The raw dataset $\mathcal{D}_i$ and the private
component $\mathbf{v}_i$ never leave the device.

\subsection{Hierarchical Federated Aggregation}

\textbf{MEC-level aggregation.}
A MEC node $j$ collects only perturbed shared components from all
$K_j$ devices within its coverage area. It computes a weighted average
of the received shared models proportional to each device's local
sample size:
\begin{equation}
    \mathbf{w}_{s,\mathrm{region},j} =
    \frac{\displaystyle\sum_{i=1}^{K_j} n_i\,\tilde{\mathbf{w}}_{s,i}}
         {\displaystyle\sum_{i=1}^{K_j} n_i}
\end{equation}
The MEC node uses $\mathbf{w}_{s,\mathrm{region},j}$ as the updated regional
initialization distributed back to nearby phones.

\textbf{Cloud-level aggregation.}
The central cloud server receives only regional shared models from all
$J$ MEC nodes and performs a second-level weighted aggregation
proportional to the total number of device-samples contributed per MEC
node:
\begin{equation}
    \mathbf{w}_{s,\mathrm{cloud}} =
    \frac{\displaystyle\sum_{j=1}^{J} N_j\,\mathbf{w}_{s,\mathrm{region},j}}
         {\displaystyle\sum_{j=1}^{J} N_j}
\end{equation}
where $N_j = \sum_{i \in j} n_i$ is the total number of samples
aggregated under MEC node $j$. The updated global shared component is
broadcast to MEC nodes, which use it to initialize or refresh regional
models.

\textbf{Region switching.}
When a user moves from one MEC region to another, only the shared
component is replaced. For example, if a phone moves from Toulouse to
Paris, the effective model changes from
$\mathbf{w}_{s,\mathrm{Toulouse}}+\mathbf{v}_i$ to
\begin{equation}
    \mathbf{w}_i =
    \mathbf{w}_{s,\mathrm{Paris}} + \mathbf{v}_i
\end{equation}
The private vector $\mathbf{v}_i$ is preserved exactly. This prevents
loss of personalization while allowing the user to benefit from the new
regional MEC model.

\begin{algorithm}[!t]
\caption{Region switching while preserving personalization}
\label{alg:region_switching}
\begin{algorithmic}[1]
\STATE \textbf{Input:} current region $r$, detected region $r'$,
current shared weights $\mathbf{w}_{s,r}$, personal weights
$\mathbf{v}_i$
\IF{$r'=r$}
    \STATE Keep current model
    $\mathbf{w}_i \leftarrow \mathbf{w}_{s,r}+\mathbf{v}_i$.
\ELSE
    \STATE Request $\mathbf{w}_{s,r'}$ from the nearest MEC node.
    \IF{MEC model is unavailable}
        \STATE Request regional fallback
        $\mathbf{w}_{s,r'}$ from the cloud backend.
    \ENDIF
    \IF{no regional shared model is reachable}
        \STATE Keep $\mathbf{w}_{s,r}$ until connectivity is restored.
    \ELSE
        \STATE Replace only the shared component:
        $\mathbf{w}_s \leftarrow \mathbf{w}_{s,r'}$.
        \STATE Preserve $\mathbf{v}_i$ unchanged on the phone.
        \STATE Set effective model
        $\mathbf{w}_i \leftarrow \mathbf{w}_{s,r'}+\mathbf{v}_i$.
        \STATE Update active region marker $r \leftarrow r'$.
    \ENDIF
\ENDIF
\STATE \textbf{Output:} personalized effective model $\mathbf{w}_i$
\end{algorithmic}
\end{algorithm}

\subsection{Personalized Hierarchical FL Protocol}

The complete learning procedure implemented by the prototype is
summarized as Algorithm~\ref{alg:pfl}. The algorithm highlights the
initialization chain, the local personalization step, and the fact that
only shared components are aggregated upstream.

\begin{algorithm}[!t]
\caption{Personalized hierarchical federated learning}
\label{alg:pfl}
\begin{algorithmic}[1]
\STATE \textbf{Input:} global shared model
$\mathbf{w}_{s,\mathrm{cloud}}$, MEC region $r$, phone user $i$, local
dataset $\mathcal{D}_i$
\STATE Cloud maintains $\mathbf{w}_{s,\mathrm{cloud}}$ from regional
shared submissions.
\STATE When MEC $r$ starts, request $\mathbf{w}_{s,r}$ from the cloud.
\IF{no regional model exists}
    \STATE $\mathbf{w}_{s,r} \leftarrow
    \mathbf{w}_{s,\mathrm{cloud}}$
\ENDIF
\STATE Phone requests $\mathbf{w}_{s,r}$ from the nearest MEC, then from
the cloud fallback if the MEC is unavailable.
\STATE For a new phone user, set $\mathbf{v}_i \leftarrow \mathbf{0}$.
\STATE Compute $\mathbf{w}_i=\mathbf{w}_{s,r}+\mathbf{v}_i$ for local
inference using profile, physiology, check-in, and MEC context.
\STATE Update $\mathbf{w}_{s,r}$ and $\mathbf{v}_i$ locally with
separate learning rates on
$\nabla\mathcal{L}_i(\mathbf{w}_{s,r}+\mathbf{v}_i)$.
\STATE Transmit only $\tilde{\mathbf{w}}_{s,i}$ and $n_i$ to the MEC.
\STATE MEC validates the update schema and rejects explicit
$\mathbf{v}_i$ or raw-data fields.
\STATE MEC computes $\mathbf{w}_{s,\mathrm{region},r} =
\sum_i(n_i/N)\tilde{\mathbf{w}}_{s,i}$.
\STATE MEC uploads only the regional shared model and aggregate metadata.
\STATE On region switch, replace $\mathbf{w}_{s,r}$ with
$\mathbf{w}_{s,r'}$ and preserve $\mathbf{v}_i$.
\end{algorithmic}
\end{algorithm}

\subsection{Profile Vector Obfuscation}

To allow cloud-side synchronization of the ASD profile across user
devices without exposing the raw sensitivity vector, a user-specific
linear obfuscation is applied before any data leaves the device. A
transformation matrix $\mathbf{M}_u \in \mathbb{R}^{n \times n}$ is
constructed deterministically from key material derived from the user's
authentication credentials via a Key Derivation Function. The stored
representation of the profile is
\begin{equation}
    \mathbf{y} = \mathbf{M}_u \cdot \mathbf{x}
\end{equation}
The matrix $\mathbf{M}_u$ is never persisted; it is recomputed on
demand from the credentials at authentication time. Profile
reconstruction on a new device proceeds as
\begin{equation}
    \mathbf{x} = \mathbf{M}_u^{-1} \cdot \mathbf{y}
\end{equation}
For this reconstruction to be well-defined, $\mathbf{M}_u$ must be
invertible. This is enforced by construction: the KDF output seeds a
deterministic pseudorandom generator that produces a full-rank matrix.
The obfuscated vector $\mathbf{y}$ is subsequently encrypted before
storage as described in Section~\ref{sec:crypto}. This matrix
transformation is therefore not intended to replace encryption, but to
act as an additional obfuscation layer. If an attacker observes a stored
numeric vector or an intermediate serialized representation, the values
do not directly correspond to interpretable sensitivities such as
noise, light, crowd density, or routine disruption. Without the
credential-derived matrix, the transformed coordinates are difficult to
interpret semantically even before considering the cryptographic
protection applied at rest.

\begin{algorithm}[!t]
\caption{Profile matrix obfuscation and recovery}
\label{alg:matrix_obfuscation}
\begin{algorithmic}[1]
\STATE \textbf{Input:} normalized ASD profile $\mathbf{x}$, password,
user salt, and fresh DEK
\STATE Derive seed material
$s \leftarrow \mathrm{KDF}_{\mathrm{Argon2id}}(\mathrm{password},
\mathrm{salt})$
\STATE Use $s$ to seed a deterministic CSPRNG.
\STATE Generate candidate matrix $\mathbf{M}_u$ until it is full-rank,
or construct an invertible matrix directly.
\STATE Compute obfuscated profile $\mathbf{y} \leftarrow
\mathbf{M}_u\mathbf{x}$.
\STATE Encrypt $\mathbf{y}$ with the DEK to obtain $C_{\mathbf{y}}$.
\STATE Derive KEK from the password and encrypt the DEK as
$C_{\mathrm{DEK}}$.
\STATE Store only $C_{\mathbf{y}}$, $C_{\mathrm{DEK}}$, and salts in the
database.
\STATE \textbf{Recovery:} after authentication, recompute
$\mathbf{M}_u$ from the same KDF seed.
\STATE Decrypt $C_{\mathrm{DEK}}$, then decrypt $C_{\mathbf{y}}$.
\STATE Recover $\mathbf{x} \leftarrow \mathbf{M}_u^{-1}\mathbf{y}$ in
memory for local inference.
\STATE Discard plaintext keys, $\mathbf{M}_u$, $\mathbf{y}$, and
$\mathbf{x}$ when the session no longer needs them.
\end{algorithmic}
\end{algorithm}

\subsection{Cryptographic Data Protection}
\label{sec:crypto}

The system employs a two-layer key hierarchy to protect stored data. A
Data Encryption Key (DEK) is generated per user as a cryptographically
random symmetric key. The obfuscated profile vector $\mathbf{y}$ is
encrypted under AES-256-GCM using the DEK:
\begin{equation}
    C_{\mathbf{y}} =
    \mathrm{Enc}_{\mathrm{AES\text{-}256\text{-}GCM}}(\mathrm{DEK},\;\mathbf{y})
\end{equation}
The DEK is itself encrypted under a Key Encryption Key (KEK) derived
from the user's password and a random salt via Argon2id:
\begin{equation}
    \mathrm{KEK} = \mathrm{KDF}_{\mathrm{Argon2id}}(\mathrm{password},\;\mathrm{salt})
\end{equation}
\begin{equation}
    C_{\mathrm{DEK}} =
    \mathrm{Enc}_{\mathrm{AES\text{-}256\text{-}GCM}}(\mathrm{KEK},\;\mathrm{DEK})
\end{equation}
Only $C_{\mathbf{y}}$ and $C_{\mathrm{DEK}}$ are stored in the
database alongside the salt. The password and plaintext DEK are never
persisted. At authentication, the KEK is recomputed from the submitted
password and the stored salt, the DEK is decrypted transiently in
memory, and the profile is decrypted for use. On devices equipped with
hardware-backed secure storage such as Apple Secure Enclave, Android
StrongBox, or a Trusted Platform Module, the KEK or DEK may
alternatively be bound to hardware-attested key material, providing an
additional layer of protection against software-level extraction.

This mechanism supports cross-device recovery of the questionnaire
profile but not automatic recovery of all phone-local learning state.
When a user authenticates on a new device, the encrypted database record
can be used to recover the obfuscated and encrypted ASD survey profile,
provided that the user supplies the correct credentials. The personal
model component $\mathbf{v}_i$, however, is designed as a phone-local
adaptation vector and is not synchronized in the current prototype. This
choice protects personalization from server-side exposure but creates a
portability limitation: a new device can reconstruct the user's survey
profile and initialize $\mathbf{w}_s$ from the MEC or cloud, but it
starts with a fresh local personal vector unless a future secure
transfer mechanism is added.

One possible extension is to apply the same protection pattern to
$\mathbf{v}_i$ when cross-device transfer is explicitly requested by the
user. In such a design, the personal vector could be transformed with a
credential-derived matrix, encrypted with a DEK, and stored as an
optional encrypted recovery object. This would preserve the principle
that the server cannot interpret the personalized component while giving
users a controlled way to migrate local adaptation between trusted
devices.


% ---------------------------------------------------------------
\section{Implementation}
% ---------------------------------------------------------------

\subsection{Implemented Components}

The prototype is organized into four implementation modules that map
directly to the architecture. The \texttt{backend} module is the cloud
layer, implemented with TypeScript, Express, Prisma, and SQLite. It
handles authentication, encrypted profile synchronization, derived
check-in storage, admin views, MEC registration, and global/regional
model endpoints. The \texttt{ui} module is a Python/NiceGUI web
application that simulates the phone client and executes local
personalized inference. The \texttt{mec} module is a Docker Compose
stack containing Mosquitto, a FastAPI MEC API, a context service, and a
regional federated aggregation service. The \texttt{ml} module contains
a PyTorch/Flower prototype used to validate the personalized FL logic
independently from the running web interface.

\subsection{Startup and Initialization Sequence}

The root launcher \texttt{run\_project.ps1} starts the backend, the UI,
and the MEC stack when Docker is available. The first component that
must be available is the cloud backend, because it owns the global
shared model and encrypted account records. It exposes
\texttt{/api/federated/global-model},
\texttt{/api/federated/regional-model}, and
\texttt{/api/federated/updates}. These endpoints are used only for
shared model coordination; they do not receive raw survey answers,
physiological streams, smartwatch data, or the private personal
component $\mathbf{v}_i$.

When a MEC node starts, its FastAPI service registers the region and
requests a regional shared model from the backend. If a model already
exists for that region, the MEC uses it. If no regional state exists,
the MEC initializes its regional model from the cloud global model.
After initialization, the MEC exposes \texttt{/fl/model} to nearby phone
clients and starts listening for phone-to-MEC shared updates on MQTT.
The Mosquitto broker routes low-latency context and update messages,
while the \texttt{mec-fl} service performs regional weighted averaging.

When a phone session starts, the UI authenticates with the backend and
recovers the encrypted survey profile if one exists. The profile is
decrypted and de-obfuscated only for the authenticated session. The
phone then requests $\mathbf{w}_s$ from the nearest MEC through
\texttt{/fl/model}. If the MEC is unreachable, the phone requests the
regional shared model from the cloud backend. If both are unavailable,
the local service uses a deterministic default so that discomfort
prediction remains available offline. For a new user, the personal
component $\mathbf{v}_i$ is initialized to zero.

\subsection{Runtime Phone Workflow}

During use, the phone-side application performs four local operations.
First, onboarding converts the sensory and behavioral survey into the
profile vector $\mathbf{x}$. Second, check-ins collect current internal
state, local context, and optional wearable signals. In a smartphone
deployment, BPM, heart-rate variability, activity level, sleep quality,
or similar data can be read from a paired smartwatch with user
permission. Third, the phone obtains regional situational context from
the MEC, such as metro delay, platform crowding, construction noise, or
quiet-area availability. Fourth, the local model evaluates
$\mathbf{w}_i=\mathbf{w}_s+\mathbf{v}_i$ on the combined feature vector
and produces a discomfort estimate.

After a check-in, the phone may perform a local personalization step.
Both $\mathbf{w}_s$ and $\mathbf{v}_i$ are updated locally, but only the
shared component is prepared for communication. The service serializes a
payload containing shared weights, sample count, region identifier, and
coarse metadata. It excludes raw check-in values, raw wearable streams,
the ASD profile vector, and the personal component $\mathbf{v}_i$.
Smartwatch integration therefore enriches local inference without
changing the external privacy boundary.

\subsection{MEC Runtime and Regional Context}

At runtime, the MEC performs two jobs. Its context services collect or
simulate regional mobility conditions and publish them on
\texttt{tsa/context/*}. In the prototype, a Station Controller simulates
noise level, crowd density, and service delay because physical station
sensors were not available. Phone clients subscribe to the context
topics relevant to their current region and append these values locally
before inference.

The MEC also performs regional aggregation. Phones publish shared
updates on \texttt{tsa/fl/update}. The \texttt{mec-fl} service validates
the update schema, accepting fields such as \texttt{shared\_weights} or
\texttt{sharedWeights} and rejecting payloads that explicitly contain
personal-component or raw-data fields. This is protocol enforcement, not
a proof against malicious encoding inside nominal shared weights. The
received shared model vectors are reconstructed from JSON lists and
aggregated using the weighted FedAvg computation defined in~(10). The
resulting regional shared model is exposed again through
\texttt{/fl/model} for phone initialization and refresh.

\subsection{Cloud Runtime and Synchronization}

The backend receives only regional shared models and aggregate metadata
from MEC nodes through \texttt{/api/federated/updates}. It uses these
regional submissions to update the global shared model and exposes the
new version for downstream MEC initialization. Separately, it manages
encrypted profile synchronization. Authentication uses Express sessions:
the backend verifies the password hash, reconstructs credential-derived
material, decrypts the DEK, and recovers the transformed profile for the
authenticated session. The database stores encrypted profile material,
encrypted DEKs, salts, derived check-in metrics, MEC metadata, regional
shared model states, and the global shared model state. It does not
store plaintext ASD profiles or phone-local personal model components.

\subsection{Standalone Federated Learning Validation}

The \texttt{ml} module validates the learning logic outside the live
MEC stack. It uses PyTorch and Flower with simulated clients. Each
client keeps a private $\mathbf{v}_i$, computes gradients on
$\mathbf{w}_s+\mathbf{v}_i$, updates both local components, and returns
only the protected shared component. The server coordinates aggregation
rounds using FedAvg and updates the global shared model without
accessing any local dataset. This standalone simulation is not a
replacement for the MEC runtime; it is a controlled validation of the
personalized FL update rule, aggregation behavior, and privacy boundary.

\begin{table*}[!t]
    \centering
    \caption{Main Communication Channels and Privacy Role\label{tab:protocol}}
    \scriptsize
    \renewcommand{\arraystretch}{1.2}
    \begin{tabularx}{\textwidth}{
        >{\raggedright\bfseries\arraybackslash}p{3.6cm}
        >{\raggedright\arraybackslash}p{3.0cm}
        >{\raggedright}X
        >{\raggedright\arraybackslash}X}
        \toprule
        \rowcolor{gray!15}
        \textbf{Channel} & \textbf{Direction} &
        \textbf{Payload} & \textbf{Privacy Role} \\
        \midrule
        \texttt{/api/federated/global-model} &
        Cloud $\rightarrow$ MEC &
        Global shared weights and version &
        Initializes new MEC nodes without user data \\
        \midrule
        \texttt{/api/federated/regional-model} &
        Cloud $\rightarrow$ MEC/Phone &
        Regional shared model or cloud fallback &
        Provides fallback when MEC is unreachable \\
        \midrule
        \texttt{/fl/model} &
        MEC $\rightarrow$ Phone &
        Regional shared component $\mathbf{w}_s$ &
        Initializes phone shared model only \\
        \midrule
        \texttt{tsa/fl/update} &
        Phone $\rightarrow$ MEC &
        Shared weights and sample count &
        Excludes $\mathbf{v}_i$ and raw data \\
        \midrule
        \texttt{tsa/context/*} &
        MEC $\leftrightarrow$ Phone &
        Situational mobility context &
        Adds local context without exposing profile \\
        \midrule
        \texttt{/api/federated/updates} &
        MEC $\rightarrow$ Cloud &
        Regional shared model and aggregate metadata &
        Enables cloud aggregation without phone records \\
        \bottomrule
    \end{tabularx}
\end{table*}


% ---------------------------------------------------------------
\section{Evaluation and Results}

\subsection{Evaluation Criteria}

The prototype evaluation focuses on technical feasibility rather than
clinical efficacy. Four criteria are considered. First, the local
inference path must remain available even when the MEC is unreachable,
because the user may still need support during connectivity loss.
Second, the federated path must enforce the privacy boundary by
transmitting only shared model components and aggregate metadata. Third,
regional adaptation must be observable: a MEC node should initialize
from the cloud, accept nearby shared updates, and expose a regional
model to phones. Fourth, personalization must survive region switching,
meaning that a new regional shared component can be loaded without
overwriting the private vector $\mathbf{v}_i$.

These criteria correspond to the main claims of the architecture. They
also define the next empirical protocol. Once participant data are
available, the system should be compared against a cloud-only baseline,
a MEC-only shared model, a purely local model, and the proposed
personalized hierarchical model. Metrics should include prediction
accuracy, calibration of discomfort scores, latency, communication
volume, battery impact, and subjective usability. Privacy evaluation
should report the type of data visible at each tier, the amount of noise
added to shared updates, and the resulting privacy budget when a formal
accountant is integrated.

\subsection{Federated Learning Prototype Evaluation}

The federated learning prototype was evaluated locally using Flower's
simulation runtime. The objective of this evaluation was not to measure
final predictive performance on a real ASD mobility dataset, which is yet to be done, but to validate the execution of the complete
privacy-preserving training pipeline. The experiment verified that
simulated phone clients can train locally, protect their model updates,
transmit only perturbed parameters, and contribute to a global FedAvg
aggregation round.

Due to local memory constraints in the WSL execution environment, the
simulation was configured with two virtual supernodes. At each round,
one client was sampled for local training and evaluation. The server was
configured for three federated rounds, one local epoch per client, a
clipping norm of $C=1.0$, and Gaussian update noise with standard
deviation $\sigma=0.01$. The round timeout was increased to 120 seconds
to account for the overhead of the Flower and Ray simulation backend.

\begin{table}[!t]
    \centering
    \caption{Federated Learning Simulation Results Over Three Rounds\label{tab:fl_results}}
    \scriptsize
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{ccc}
        \toprule
        \textbf{Round} & \textbf{Distributed Loss} & \textbf{Client Failures} \\
        \midrule
        1 & 0.7357 & 0 \\
        2 & 0.6883 & 0 \\
        3 & 0.6765 & 0 \\
        \bottomrule
    \end{tabular}
\end{table}

As shown in Table~\ref{tab:fl_results}, the distributed loss decreases
across the three rounds, from 0.7357 to 0.6765. This confirms that the
global model receives useful information from local client updates and
that the FedAvg aggregation loop executes correctly. No client failures
were observed in the final stabilized configuration. The experiment
therefore validates the functional feasibility of the federated
learning workflow in the proposed architecture.

The evaluation also confirmed the privacy-preserving data flow of the
prototype. Raw synthetic user records remained local to each simulated
phone client. The server received only clipped and noised model
parameters, together with coarse training metrics. This behavior is
consistent with the system objective of avoiding centralised exposure
of sensitive ASD-related mobility profiles.

The loss values should be interpreted as an engineering validation
signal rather than as a clinical performance claim. The goal of this
experiment was to verify that the training loop, parameter exchange,
aggregation, and privacy filters operate coherently. A future empirical
study will require participant-approved labels, repeated mobility
episodes, and comparison against baselines such as a non-personalized
global model, a regional-only model, and a purely local model. These
baselines will make it possible to quantify the contribution of
personalization, regional MEC context, and hierarchical aggregation.

\subsection{End-to-End Prototype Validation}

The implemented prototype was also validated as an integrated
Cloud--MEC--Device stack. The objective was to verify that each system
component executes its expected role and that the privacy boundaries in
Table~\ref{tab:privacy_boundary} are respected during normal operation.
The web application executes the local personalized model during
check-in, the MEC layer exposes regional shared-model initialization,
and the backend stores cloud and regional shared-model metadata.

\begin{table}[!t]
    \centering
    \caption{Functional Validation of the Implemented Prototype\label{tab:functional_validation}}
    \scriptsize
    \renewcommand{\arraystretch}{1.2}
    \begin{tabularx}{\columnwidth}{
        >{\raggedright\bfseries}p{2.0cm}
        >{\raggedright}X
        >{\raggedright\arraybackslash}p{1.7cm}}
        \toprule
        \rowcolor{gray!15}
        \textbf{Function} & \textbf{Validated Behavior} & \textbf{Result} \\
        \midrule
        Phone-local AI &
        Check-in features are evaluated by the personalized
        $\mathbf{w}_s+\mathbf{v}_i$ model in the web application &
        Functional \\
        \midrule
        Personalization privacy &
        The phone stores $\mathbf{v}_i$ locally and publishes only
        shared weights to \texttt{tsa/fl/update} &
        Functional \\
        \midrule
        MEC aggregation &
        The MEC aggregator accepts shared weights and rejects personal
        component fields &
        Functional \\
        \midrule
        Cloud coordination &
        The backend exposes global/regional initialization endpoints and
        accepts regional shared updates &
        Functional \\
        \midrule
        Region switching &
        The phone replaces only $\mathbf{w}_s$ when a new region model
        is received and preserves $\mathbf{v}_i$ &
        Functional \\
        \midrule
        Context enrichment &
        MEC situational context can be injected into local inference
        without transmitting the user's ASD profile &
        Functional \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{Prototype Data and Demonstration Results}

Because no real participant study has yet been conducted, the current
data are prototype data generated by scripted phone clients, manual web
check-ins, and MEC environment simulations. The synthetic check-in,
context, and model-update examples used for these tests were generated
with AI assistance and then manually reviewed to ensure that they
matched the expected feature structure of the prototype. These data
exercise the same communication and storage paths intended for
deployment: check-ins produce local discomfort predictions, phone
clients generate shared model updates, MEC nodes aggregate regional
shared models, and the cloud backend records global/regional model
state. This provides evidence of technical feasibility, but it should
not be interpreted as clinical validation of ASD stress prediction.

\begin{table}[!t]
    \centering
    \caption{Prototype Demonstration Results Using Synthetic and Manual Test Data\label{tab:demo_results}}
    \scriptsize
    \renewcommand{\arraystretch}{1.2}
    \begin{tabularx}{\columnwidth}{
        >{\raggedright\bfseries}p{2.2cm}
        >{\raggedright}X
        >{\raggedright\arraybackslash}p{1.5cm}}
        \toprule
        \rowcolor{gray!15}
        \textbf{Scenario} & \textbf{Observation} & \textbf{Outcome} \\
        \midrule
        Normal check-in &
        The web app generates a local stress estimate and logs derived
        metrics rather than raw context &
        Passed \\
        \midrule
        Delayed metro context &
        MEC situational delay is represented as local contextual input,
        increasing the model's available environmental evidence &
        Passed \\
        \midrule
        Shared update upload &
        Phone clients publish shared weights and sample counts; no
        personal component is included &
        Passed \\
        \midrule
        Regional aggregation &
        MEC computes weighted FedAvg over shared updates and republishes
        a regional shared model &
        Passed \\
        \midrule
        Cloud synchronization &
        Backend accepts regional shared model metadata and updates the
        global shared model version &
        Passed \\
        \bottomrule
    \end{tabularx}
\end{table}

% ---------------------------------------------------------------
\section{Discussion}
\subsection{Limitations of the Federated Learning Prototype}

The current federated learning evaluation remains preliminary. First,
the dataset used in the prototype is synthetic and only approximates the
structure of future ASD mobility features. It is sufficient to validate
the software pipeline, but not to draw conclusions about real-world
prediction accuracy. Second, differential privacy is treated as an
optional update-hardening mechanism rather than a core requirement of
the architecture; no formal privacy budget $(\varepsilon, \delta)$ has
yet been calibrated. Third, the current evaluation focuses on
functional correctness and architectural feasibility. A larger benchmark
should next quantify the benefit of the full Cloud--MEC--Device
hierarchy against cloud-only, MEC-only, and purely local baselines under
different mobility regions and network conditions.

Despite these limitations, the prototype demonstrates that the proposed
architecture can support local training, protected update transmission,
and server-side aggregation without collecting raw user data.

A second limitation concerns device portability. The encrypted database
record allows an authenticated user to recover the ASD survey profile on
a new device, because the profile is stored as an encrypted and
matrix-obfuscated vector. The personalized model component
$\mathbf{v}_i$ is different: by design it remains local to the original
phone and is not uploaded to the MEC or cloud. Consequently, a new
device can initialize from the current MEC or cloud shared component and
recover the survey profile, but it cannot automatically recover the
previous device's learned personal adaptation. A future version could
make this optional by applying the same obfuscation and encryption
mechanisms to $\mathbf{v}_i$ before storing or transferring it, while
still ensuring that the server cannot interpret the personal vector.

\subsection{Security and Privacy Limitations}

The matrix transformation should be interpreted as an additional
obfuscation mechanism rather than a standalone cryptographic guarantee.
Encryption remains the primary confidentiality mechanism for stored
profile data. Similarly, differential privacy should be understood as an
optional protection for federated shared updates, not as the foundation
of privacy in the current design. The current prototype also does not
yet implement secure aggregation, Byzantine-robust aggregation, formal
client attestation, or a calibrated differential privacy accountant.
MQTT is used for local prototype communication and should be deployed
with TLS, authentication, and topic-level access control in a production
setting. Finally, the prototype data are synthetic or manually
generated; real-world validation requires an ethically approved study
with ASD participants or caregivers, carefully designed consent
materials, and controlled evaluation of prediction accuracy, latency,
usability, and safety.

% ---------------------------------------------------------------
\section{Conclusion}
% ---------------------------------------------------------------

This paper presented a privacy-preserving intelligent mobility
prototype for individuals with ASD. The system combines local
discomfort prediction, encrypted and matrix-obfuscated profile
synchronization, MEC-based situational context, and hierarchical
personalized federated learning. Its main design principle is that raw
sensory profiles, physiological signals, mobility context, and the
personal adaptation component of the model remain on the phone, while
only shared model updates are aggregated by MEC nodes and then by the
cloud.

The implemented prototype demonstrates the feasibility of this
Cloud--MEC--Device architecture. The web application performs local AI
inference using the effective model
$\mathbf{w}_i=\mathbf{w}_s+\mathbf{v}_i$, MEC nodes provide regional
context and aggregate shared updates, and the cloud maintains global and
regional shared model initialization. Region switching is also handled
without losing personalization, because only the shared component is
replaced while the private component is preserved. Prototype validation
with synthetic and manual demonstration data confirmed that the main
software paths execute as intended and that no raw personal data is
required for federated learning.

Future work will focus on training and evaluating the model with
ethically collected ASD mobility data, calibrating formal differential
privacy guarantees, strengthening secure aggregation and client
attestation, and measuring prediction accuracy, latency, usability, and
perceived safety in realistic mobility scenarios. Another practical
extension is deeper smartwatch integration, including continuous BPM,
heart-rate variability, activity, sleep, and stress proxy signals, while
preserving the same local-only processing boundary for raw wearable
data.


\section*{Acknowledgment}

The authors acknowledge the use of OpenAI ChatGPT and OpenAI Codex for
formal English formulation, grammar, clarity, and LaTeX presentation
checks. These tools were also used to generate synthetic prototype test
examples for check-ins, context values, and model-update payloads; these
examples were manually reviewed and used only to exercise the software
pipeline. The authors evaluated all suggestions and generated examples
before accepting or modifying them. All corrections and wording changes
retained in the final manuscript faithfully reflect the authors'
original work, technical design, and interpretation. No aspect of the
research problem, system architecture, implementation, analysis, or
scientific interpretation of this report was generated by an AI tool.


% ---------------------------------------------------------------
\bibliographystyle{IEEEtran}
\begin{thebibliography}{1}

\bibitem{maclennan2023}
K.~MacLennan et al., ``It is a big spider web of things: Sensory
experiences of autistic adults in public spaces,'' \textit{Autism in
Adulthood}, vol.~5, no.~4, pp.~411--422, 2023, doi:
10.1089/aut.2022.0024.

\bibitem{cena2023}
F.~Cena, N.~Mauro, and A.~Rapp, ``How do sensory features of places
impact on spatial exploration of people with autism? A user study,''
\textit{Inf. Technol. Tourism}, vol.~25, no.~1,
pp.~105--132, 2023, doi: 10.1007/s40558-023-00244-1.

\bibitem{garroussi2025}
H.~Garroussi et al., ``A systematic review of data privacy in
Mobility-as-a-Service (MaaS),'' \textit{Transp. Res. Interdiscip.
Perspect.}, vol.~31, Art.~no.~101254, 2025, doi:
10.1016/j.trip.2024.101254.

\bibitem{beck2025}
K.~B.~Beck et al., ``Guidelines for the creation of accessible consent
materials and procedures: Lessons from research with autistic people and
people with intellectual disability,'' \textit{Autism in Adulthood},
early access, 2025, doi: 10.1089/aut.2024.0263.

\bibitem{adnan2022}
M.~Adnan, S.~Kalra, J.~C.~Cresswell, G.~W.~Taylor, and H.~R.~Tizhoosh,
``Federated learning and differential privacy for medical image
analysis,'' \textit{Sci. Rep.}, vol.~12, Art.~no.~1953, 2022,
doi: 10.1038/s41598-022-05539-7.

\bibitem{alwarafy2020}
A.~Alwarafy, K.~A.~Al-Thelaya, M.~Abdallah, J.~Schneider, and
M.~Hamdi, ``A survey on security and privacy issues in edge-computing-
assisted Internet of Things,'' \textit{IEEE Internet Things J.},
vol.~8, no.~6, pp.~4004--4022, 2021, doi: 10.1109/JIOT.2020.3015432.

\end{thebibliography}

\end{document}
